Avoiding tipping points in fisheries management: A Bayesian Non-Parametric approach

### 2014-06-03 manuscripts/manuscript.tex

In order to find that policy, we introduce the value function$V_t(x_t)$
representing the total discounted catch from time $t$ onwards given that
$X_t=x$. This value function satisfies an equation of Stochastic Dynamic
Programming (SDP, Mangel and Clark 1988, Clark and Mangel 2000, Clark
2009), sometimes called the Bellman-Hamilton-Jacobi equation (Mangel
2014).

\begin{equation}
V_t(x) = \max_{h_t} \left[R(h_t, x_t) + \delta \cdot \mathbf{E}_{X_t+1} V_{t+1}( Z_{t+1} f(x - h_t)) \right]
\end{equation}

That is, at time $t$, when population size is $x$ and harvest $h_t$ is
applied, the immediate return is $R(h_t, x_t)$ and the population size
after the take is $x-h_t$, which is then translated into $X_{t+1}$ by
Eqn 1. We consider the finite time problem with \$T=\$1000, which we
solve using value iteration algorithm (Mangel and Clark 1988, Clark and
Mangel 2000).

@@

\[y = f(x) + \varepsilon\]

$\varepsilon$ IID normal, variance $\sigma_n^2$. Then under the GP, the
predicted data $X_p$ given the observations obeys

@@

\% latex table generated in R 3.1.0 by xtable 1.7-3 package \% Mon Jun 2
00:55:24 2014

>>>>

In order to find that policy, we introduce the value function$V_t(x_t)$
representing the total discounted catch from time $t$ onwards given that
$X_t=x$. This value function satisfies an equation of Stochastic Dynamic
Programming (SDP, Mangel and Clark 1988, Clark and Mangel 2000, Clark
2009), sometimes called the Bellman-Hamilton-Jacobi equation (Mangel
2014),

\begin{equation}
V_t(X_t) = \max_{h_t} \left{ R(h_t, X_t) + \delta \cdot \mathbf{\mathrm{E}}_{X_t+1} \left[ V_{t+1}( X_{t+1}) | X_t, h_t \right] \right}
\end{equation}

where expectation is taken over all possible values of the next state,
$X_t{+1}$, and maximized over all possible choices of harvest, $h_t$.
That is, at time $t$, when population size is $x$ and harvest $h_t$ is
applied, the immediate return is $R(h_t, X_t)$. In the classical case,
the sole source of uncertainty is the process noise term, $Z$, and thus
the expectation above is equivalent to taking expectations over $Z$.
That is

\[ \mathbf{\mathrm{E}}_{X_t+1} \left[ V_{t+1}( X_{t+1}) | X_t, h_t \right] = \mathbf{\mathrm{E}}_{Z} \left[ V_{t+1}( Z f(X_t - h_t))  | X_t, h_t \right] \]

where the population size after the take is $X_t-h_t$, which is then
translated into $X_{t+1}$ by Eqn 1.

In a Bayesian decision framework, the parameters governing the dynamics
are also uncertain, so the expectation in (1) involves averaging over
the posterior distribution for the parameters, as well. That is,

\[\mathbf{\mathrm{E}}_{X_t+1} \left[ V_{t+1}( X_{t+1}) | X_t, h_t \right] = \mathbf{\mathrm{E}}_{\theta,\mathrm{data}} \{ \mathbf{\mathrm{E}}_{Z | \theta, \mathrm{data}} \left[ V_{t+1}( Z f(X_t - h_t))  | X_t, h_t \right] \}\]

In the nonparametric case, the function $f$ too is uncertain and the
expectation for the next state includes uncertainty in $f$ as well. That
is

\[\mathbf{\mathrm{E}}_{X_t+1} \left[ V_{t+1}( X_{t+1}) | X_t, h_t \right] = \mathbf{\mathrm{E}}_{\theta,\mathrm{data}} \{ \mathbf{\mathrm{E}}_{f, Z | \theta, \mathrm{data}} \left[ V_{t+1}( Z f(X_t - h_t))  | X_t, h_t \right] \}\]

We consider the finite time problem with \$T=\$1000, which we solve
using value iteration algorithm (Mangel and Clark 1988, Clark and Mangel
2000).

@@

\[X_{t+1} = f(X_t) + \varepsilon\]

$\varepsilon$ IID normal, variance $\sigma_n^2$. Note that we have
chosen to assume additive noise. While we could just as easily consider
log-normal noise as in the parametric models, we make this choice to
bring home the point that the Gaussian process approach need not have
structurally correct noise form either.

Then under the GP, given a vector of observed data $X_0$ the predicted
data $X_p$ obeys

@@

\% latex table generated in R 3.1.0 by xtable 1.7-3 package \% Mon Jun 2
17:34:16 2014
